On the uniqueness and stability of dictionaries for sparse representation of noisy signals

### 2017-08-17 manuscripts/PNAS/chazthm_pnas.tex
chaz

Before moving on to proofs, we discuss briefly the tightness of our results for quantities: (a) the required coding error $\varepsilon(\delta_1, \delta_2)$ sufficient to guarantee uniqueness in terms of desired dictionary and sparse representation errors $\delta_1, \delta_2$, and (b) the number of required samples, $N$.  With respect to (a), even with fixed dictionary, the recovery of sparse representations up to an error that is linear in the noise is classically optimal (up to constants).  %Thus, having both sufficient error in coding for uniqueness being linear in terms of both the desired dictionary and sparse vector recovery error is optimal.  
We regard (b) as a basic open problem: how many samples determine the sparse linear coding model?  If $k$ is held fixed or if the size of the support set of reconstructing codes is known to be polynomial in $\bar m$ and $k$, then a practical (polynomial) amount of data is enough to determine the model.\footnote{In the latter case, a reexamination of the pigeonholing argument in the proof of Thm.~\ref{DeterministicUniquenessTheorem} shows that only a polynomial number of samples distributed over a polynomial number of supports suffices.} Reasons to be skeptical that a polynomial amount suffice in general, however, can be found in \cite{tillmann2014computational, tillmann2015computational}.

@@

Note that the explicit constants $C_1$ and $C_2$ are effectively computable. (The quantity $L_k$, may be computed as the smallest singular value of a certain matrix, while the quantity $\xi$ involves computing ``canonical angles'' between subspaces, which reduce again to an efficient singular value decomposition.) Because deciding $L_{k}(\mathbf{A}) > 0$ is already NP-hard \cite{tillmann2014computational}, it seems likely that without fixing a sparsity level there is no efficient computation $L_k$ in general; for fixed sparsity $k$, however, it is computable in polynomial time. Regardless, to calculate $C_2$ nonetheless requires an exponential number of queries to $\xi$. Thus, $C_1$ and $C_2$ are not necessarily polynomial-time computable.

>>>>

Before moving on to proofs, we discuss briefly the tightness of our number of required samples, $N$. We regard this as a basic open problem: how many samples determine the sparse linear coding model?  If $k$ is held fixed or if the size of the support set of reconstructing codes is known to be polynomial in $\bar m$ and $k$, then a practical (polynomial) amount of data is enough to determine the model.\footnote{In the latter case, a reexamination of the pigeonholing argument in the proof of Thm.~\ref{DeterministicUniquenessTheorem} shows that only a polynomial number of samples distributed over a polynomial number of supports suffices.} Reasons to be skeptical that a polynomial amount suffice in general, however, can be found in \cite{tillmann2014computational, tillmann2015computational}.

@@

We note that the explicit constants $C_1$ and $C_2$ are effectively computable. (The quantity $L_k$, may be computed as the smallest singular value of a certain matrix, while the quantity $\xi$ involves computing ``canonical angles'' between subspaces, which reduce again to an efficient singular value decomposition.) Because deciding $L_{k}(\mathbf{A}) > 0$ is already NP-hard \cite{tillmann2014computational}, it seems likely that without fixing a sparsity level there is no efficient computation $L_k$ in general; for fixed sparsity $k$, however, it is computable in polynomial time. Regardless, to calculate $C_2$ requires an exponential number of queries to $\xi$. Thus, $C_1$ and $C_2$ are not necessarily polynomial-time computable.
